The development of children's ability to track and predict turn structure in conversation

### 2015-11-30 paper/CasillasFrank-Anticip.tex
full draft of paper

\usepackage{graphicx}
% \usepackage{lineno}
\usepackage{amssymb}
\usepackage{wrapfig}
\usepackage{subfig}
\usepackage{tipa}
\usepackage{float}
\usepackage{longtable}

@@

\begin{abstract}
We investigate language acquisition through the lens of a fundamental conversational skill: Turn taking. Children begin developing turn-taking skills in infancy, but take several years to assimilate their growing knowledge of language into their turn-taking behavior. In two eye-tracking experiments, with children across a wide developmental range, we measured spontaneous predictions about upcoming speaker change while controlling the amount of linguistic information available. We found that children already predicted upcoming turns at age one, but that they integrated linguistic cues differently at different ages. Children under three showed a general advantage for prosody over lexicosyntax, contrary to prior findings for adults. Children two and older showed more predictive switches for questions than non-questions, but only when lexical information was available. We found no evidence that lexicosyntax alone guides turn prediction---instead, participants' performance was best overall with access to lexicosyntax and prosody together. Our results point to the importance of studying children's linguistic processing in the context of conversation.
\end{abstract}

@@

% \linenumbers

@@

Children participate in sequential coordination with their caregivers starting at three months of age---before they can rely on any linguistic cues in taking turns \citep[see, among others, ][]{bateson1975, hilbrink2015, jaffe2001, snow1977}. Of course, infant turn taking is different from adult turn taking in several ways. Infant turn taking is heavily scaffolded by caregivers, has distinct timing in comparison to adult turn taking, and lacks semantic content \citep{hilbrink2015, jaffe2001}. However, children's early, turn-structured social interactions are presumably a critical precursor to their conversational turn taking. Along these lines, early non-verbal interactions establish the protocol by which children come to use language with others. And as they acquire language, they also come to integrate it into the preverbal turn-taking systems.

@@

Conversation distinguishes itself from non-conversational turn-taking behaviors by the complexity of the turn sequencing involved. In the examples above (traffic, games, and monkeys) the set of sequence and action types is far more limited and predictable than what we find in everyday talk. For example, conversational turns come grouped into semantically-contingent sequences of action. The groups can span turn-by-turn exchanges (e.g., simple question--response, ``How are you?''--``Fine.'') or sequence-by-sequence exchanges (e.g., reciprocals, ``How are you?''--``Fine, and you?''--``Great!''). Sequences of action drive the conversation forward into the next, relevant sequences of talk (e.g., ''And you?''--``Great!''--``Why's that?''; \citealp{schegloff2007}). To take a turn, participants need to make predictions about what conversational content will be relevant next. In some cases, relevant next turns are somewhat obvious (e.g., question--response) while, in other cases, there are multiple relevant next actions to choose from, or no obvious next action at all (e.g., after a closing).

@@

Adults have a lot of information at their disposal to help make accurate predictions about upcoming turn content. Lexical, syntactic, and prosodic information (e.g., \textit{wh}- words, subject-auxiliary inversion, and list intonation) can all inform addressees about upcoming linguistic structure \citep{de-ruiter2006, duncan1972, ford1996, torreira2015}. Non-verbal cues (e.g., gaze, posture, and pointing) often appear at turn-boundaries and can sometimes act as late indicators of an upcoming speaker switch \citep{rossano2009, stivers2010}. Additionally, the sequential context of a turn can make it clear what will come next: Answers after questions; thanks or denial after compliments, et cetera \citep{schegloff2007}.

Prior work suggests that adult listeners primarily use lexicosyntactic information to accurately predict upcoming turn structure \citep{de-ruiter2006}. De Ruiter and colleagues \citeyearpar{de-ruiter2006} asked participants to listen to snippets of spontaneous conversation and to press a button whenever they anticipated that the current speaker was about to finish his or her turn. The speech snippets were controlled for the amount of linguistic information present; some were normal, but others had flattened pitch, low-pass filtered speech, or further manipulations. De Ruiter and colleagues found that, with pitch-flattened speech, the timing of participants' button responses was comparable to their timing with the full linguistic signal. But, when no lexical information was available, participants' responses were significantly earlier. The authors concluded that lexicosyntactic information\footnote{The ``lexicosyntactic'' condition only included flattened pitch and so was not exclusively lexicosyntactic---the speech would still have residual prosodic structure, including syllable duration and intensity.} was necessary and possibly sufficient for turn-end projection, while intonation was neither necessary nor sufficient. Congruent evidence comes from studies varying the predictability of lexicosyntactic and pragmatic content: Adults anticipate turn ends better when they can more accurately predict the exact words that will come next (\citealp{magyari2012}; see also \citealp{magyari2014}). They can also identify speech acts within the first word of an utterance \citep{gisladottir2015}, allowing them to start planning their response at the first moment possible \citep{bogels2015}.

The role of prosody for adult turn-prediction is still a matter of debate. De Ruiter and colleagues' (2006) experiment focused on the role of intonation, which is only a partial index of prosody. Prosodic structure is also tied closely to the syntax of an utterance, and so the two linguistic signals are difficult to control independently \citep{ford1996}. \citet*{torreira2015} used a combination of button-press and verbal responses to investigate the relationship between lexicosyntactic and prosodic cues in turn-end prediction. Critically, their stimuli were cross-spliced so that each item had full prosodic cues to accompany the lexicosyntax. Because of the splicing, they were able to create items that had syntactically-complete units with no intonational phrase boundary at the end. Participants never verbally responded or pressed the ``turn-end'' button when hearing a syntactically-complete phrase without an intonational phrase boundary. And when intonational phrase boundaries were embedded in multi-utterance turns, participants were tricked into pressing the ``turn-end'' button 29\% of the time. Their results suggest that listeners actually do rely on prosodic cues to execute a response (see also de \citet{de-ruiter2006}:525). These experimental findings corroborate other corpus and experimental work promoting a combination of cues (lexicosyntactic, prosodic, and pragmatic) as key for accurate turn-end prediction \citep{duncan1972, ford1996, hirvenkari2013}.

@@

The timing of children's responses to their caregivers' speech shows a non-linear pattern of fall-rise-fall from early infancy to middle childhood. A recent study by Hilbrink et al. (2015) finds that infants' turn timing at three months is often too early or too late: They start vocalizing in overlap on $~$40\% of their caregivers' turns, and their non-overlapped vocalizations come after an average inter-turn silent gap of 350--900 msec (adult average: 200 msec). Between four and nine months, children begin to reduce the number of turns happening in overlap while also improving on their average response latency. But then, later on, children's response latencies slow down again, peaking at average gaps of more than 1000 msec at nine months, with only very gradual improvement after that \citep{hilbrink2015}. While children's avoidance of overlap is nearly adult-like by nine months, the timing of their non-overlapped responses stays much longer than the 200 msec standard for the next few years \citep{garvey1984, ervin-tripp1979}.

@@

Keitel and colleagues \citeyearpar{keitel2013} addressed the random baseline issue in their study of 6-, 12-, 24-, and 36-month-olds. They asked participants to watch short videos of conversation, and tracked their eye movements at points of speaker change. They found that children's anticipatory gaze frequency was only greater than chance for 36-month-olds and adults. Their study was the first to focus on the role of linguistic processing in children's turn predictions. They showed their participants two types of conversation videos: One normal and one with flattened pitch (i.e., with flattened intonation contours), finding that only 36-month-olds were affected by a lack of intonation contours. The adult control group made equal numbers of anticipatory looks in the videos, with and without intonation contours, consistent with prior adult findings \citep{de-ruiter2006}. Keitel and colleagues concluded that children's ability to predict upcoming turn structure relies on their ability to comprehend the stimuli (emerging around 36 months), especially with respect to semantic access. They also suggest that intonation takes a secondary role in turn prediction, but only \textit{after} children acquire more sophisticated, adult-like language comprehension systems (sometime after 36 months).

Although the Keitel et al. \citeyearpar{keitel2013} study constitutes a substantial advance over previous work, it has its own limitations. Because these limitations directly inform our own study design, we review them in some detail. First, their estimates of baseline gaze frequency (``random'' in their terminology) were not random. Instead, they used gaze switches during ongoing speech as a baseline, during which switching is least likely to occur \citep{hirvenkari2013} and thereby maximizing their chances of finding a difference between gaze frequency at turn transitions and their baseline rate. A more conservative baseline would be to compare participants' looking behavior at turn transitions to their looking behavior during randomly-selected windows of time throughout the stimulus. We follow this conservative approach in our work.

@@

Despite these minor methodological issues, the Keitel et al. \citeyearpar{keitel2013} study still demonstrates intriguing age-based differences in children's ability to predict upcoming turn structure, and the results suggest that both semantic and intonational development \textit{do} play a role in children's looking patterns. Our current work thus takes this paradigm as our starting point.\footnote{See also \citet{casillas2012, casillas2013}.}

@@

\subsection{The Current Study}
We report here on the role of linguistic processing in children's predictions about upcoming turn structure. We focus in particular on how children use prosodic and lexicosyntactic information to make their predictions. Prior work has focused mainly on lexicosyntax and intonation, and not on prosody proper (\citealp{de-ruiter2006}; \citealp{keitel2013}, but see \citealp{torreira2015}), even though infants seem to acquire the basic rhythmic properties of the prosodic signal first \citep{mehler1988, moon1993, nazzi2003}.

In two eye-tracking experiments, we measured children's anticipatory gaze to upcoming responders while controlling for the amount of lexicosyntactic and prosodic information available. In Experiment 1, English-speaking participants viewed video clips of naturalistic conversation from several different languages. We used multiple languages to control for the presence of lexicosyntactic information while keeping prosodic and non-linguistic cues intact. The results showed minimal differences between the predictive looking behavior of preschoolers and adults. In Experiment 2, we created artificial (puppet) visual scenes, enabling us to separately control lexicosyntactic and prosodic cues in the conversational stimuli. In this more controlled paradigm, we found that children's predictive looking behavior improved from ages one to six, but that even one-year-olds made more anticipatory looks than would be expected by chance.

In both experiments children consistently looked faster to responders after hearing questions, compared to non-questions. Both prosodic and lexicosyntactic information played a role in children's predictions about turn structure, but the two information sources were used differently at different ages. Our findings overall support an account in which predictive processes for turn taking in conversation are present early, but their integration with linguistic information takes substantial practice.

@@

Each recording session began with a 20-minute warm-up period of spontaneous conversation during which the pair talked for five minutes on four topics (favorite foods, entertainment, hometown layout, and pets). Then we asked talkers to choose a new topic---one relevant to young children (e.g., riding a bike, eating breakfast)---and to improvise a dialogue on that topic. We asked them to speak as if they were on a children's television show in order to elicit child-directed speech toward each other. We recorded until the talkers achieved at least 30 seconds of uninterrupted discourse with enthusiastic, child-directed speech. Most talker pairs took less than five minutes to complete the task, usually by agreeing on a rough script at the start. We encouraged talkers to ask at least a few questions to each other during the improvisation. The resulting conversations were therefore not entirely spontaneous, but were as close as possible while still remaining child-oriented in topic, prosodic pattern, and lexicosyntactic construction.\footnote{All of the non-English talkers were fluent in English as a second language, and some fluently spoke a third or more language. We chose male-female pairs as a natural way of creating contrast between the two talker voices.}

%% PART UNFINISHED: link for example video

@@

Because the conversational stimuli were recorded semi-spontaneously, the duration of turn transitions and the number of speaker transitions in each video was variable. We measured the duration of each turn transition from the audio recording associated with each video. We excluded turn transitions longer than 550 msec and shorter than 90 msec, including overlapped transitions, from analysis.\footnote{Overlap occurs when a responder begins a new turn before the current turn is finished. When overlap occurs, observers cannot switch their gaze in anticipation of the response because the response began earlier than expected; participants expect conversations to proceed with ``one speaker at a time'' \citep{sacks1974}. As such, they would still be fixated on the prior speaker when the overlap started, and then would have to switch their gaze \textit{reactively} to the responder.} This left approximately equal numbers of turn transitions available for analysis in the English (N$=$20) and non-English (N$=$16) videos. On average, the inter-turn gaps for English videos (mean$=$318, median$=$302, stdv$=$112 msec) were slightly longer than for non-English videos (mean$=$286, median$=$251, stdv$=$122 msec). The longer gaps in the English videos could give them a slight advantage: Our definition of an ``anticipatory gaze shift'' includes shifts that are initiated during the gap between turns (Figure \ref{fig:criterion}), so participants had slightly more time to make anticipatory shifts in the English videos.

Questions made up exactly half of the turn transitions in the English (N$=$10) and non-English (N$=$8) videos. In the English videos, inter-turn gaps were slightly shorter for questions (mean$=$310, median$=$293, stdev$=$112 msec) than non-questions (mean$=$325, median$=$315, stdv$=$118 msec). Non-English videos did not show a large difference in transition time for questions (mean$=$270, median$=$257, stdv$=$116 msec) and non-questions (mean$=$302, median$=$252, stdv$=$134 msec).

@@

\begin{table}[h!]
\begin{center}
  \begin{tabular}{lcccl}
  \textbf{\textit{Children}} &&&& \\
    \hline
                   & Estimate & Std. Error & z value & Pr($>$$|$z$|$)    \\
    \hline
    (Intercept)   &      -0.96146 &   0.84901 & -1.132 & 0.257446 \\    
    Age             &    -0.18268 &   0.17507 & -1.043 & 0.296725   \\  
    LgCond$=$\textit{non-English} &          -3.29347 &   0.96045 & -3.429 & 0.000606 *** \\
    Type$=$\textit{non-Question}          &     -1.10129 &   0.86494 & -1.273 & 0.202925     \\
    Duration      &       3.40169 &   1.22826 &  2.770 & 0.005614  **\\
    Age*LgCond$=$\textit{non-English} &       0.52065 &   0.21190 &  2.457 & 0.014008 ** \\  
    Age*TypeS$=$\textit{non-Question}          & -0.01628 &   0.19437 & -0.084 & 0.933232     \\
    LgCond$=$\textit{non-English}* &     2.68166 &   1.35016 &  1.986 & 0.047013 * \\
    Type$=$\textit{non-Question} &&&& \\
    Age*LgCond$=$\textit{non-English}* & -0.45632 &   0.30163 & -1.513 & 0.130315  \\
    Type$=$\textit{non-Question} &&&& \\
    \hline
  &&&& \\
  \textbf{\textit{Adults}} &&&& \\
    \hline
                       &  Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\    
    \hline
    (Intercept)       &        -0.1966 &    0.6942 & -0.283 & 0.776988     \\
    LgCond$=$\textit{non-English}    &             -0.8812 &    0.9602 & -0.918 & 0.358754   \\  
    Type$=$\textit{non-Question}             &        -4.4953 &    1.3139 & -3.421 & 0.000623 *** \\
    Duration          &        -1.1227 &    1.9880 & -0.565 & 0.572238     \\
    LgCond$=$\textit{non-English}* &           3.2972 &    1.6101 &  2.048 & 0.040581 * \\ 
    Type$=$\textit{non-Question} &&&& \\
    LgCond$=$\textit{non-English}* &        1.3626 &    3.0077 &  0.453 & 0.650527     \\
    Duration &&&& \\
    Type$=$\textit{non-Question}*          &  10.5107 &    3.3459 &  3.141 & 0.001682  ** \\
    Duration &&&& \\
    LgCond$=$\textit{non-English}* & -6.3156 &    4.4926 & -1.406 & 0.159790 \\    
    Type$=$\textit{non-Question}* &&&& \\
    Duration &&&& \\
    \hline
  \end{tabular}
\end{center}
  \caption{Model output for children and adults' anticipatory gaze switches.}
\label{tab:E1-models}
\end{table}

@@

% PART UNFINISHED (potentially)
% We excluded the output of random-permutation models that did not converge to remove unreliable model results from our percentile calculations below (XX\% and XX\% of models for children and adults, respectively).

Our baseline analyses revealed that none of the significant predictors from models of the original, turn-related data can be explained by random looking. The children's data showed strong evidence of differentiation from the randomly permuted data for all four significant effects in the original model (Table \ref{tab:E1-models}: Children):  the original \textit{t}-values for language condition, gap duration, the age-language condition interaction, and the language condition-transition type interaction were all greater than 95\% of \textit{t}-values for the randomly permuted data (99.9\%, 95.5\%, 99.4\%, and 96\%, respectively; Figure \ref{fig:E1-ChiTs}). Similarly, the adults' data showed significant differentiation from the randomly permuted data for two of the three originally significant predictors---transition type and the transition type-gap duration interaction (greater than 99.9\% and 99.7\% of random \textit{t}-values, respectively)---with marginal differentiation for the interaction of language condition and transition type (greater than 94.7\% of random \textit{t}-values; Figure \ref{fig:E1-AduTs}). The effects of language condition and transition type for the real and randomly permuted data can also be observed in Figure \ref{fig:E1-randvsreal}.

@@

% PART UNFINISHED: can add a link at the end of the prior paragraph to video examples

@@

% PART UNFINISHED: update gap information
There were 85 turn transitions for analysis after excluding transitions longer than 550 msec and shorter than 90 msec. The remaining turn transitions had slightly more questions than non-question (N$=$50 and N$=$35, respectively), with transitions distributed somewhat evenly across conditions (keeping in mind that there were three \textit{normal} videos and only one limited cue video for each experiment version): \textit{Normal} (N$=$36), \textit{words only} (N$=$13), \textit{prosody only} (N$=$17), and {no speech} (N$=$19). Inter-turn gaps for questions (mean$=$365, median$=$427) were longer than those for non-questions (mean$=$302, median$=$323) on average, but gap duration was overall comparable across conditions: \textit{Normal} (mean$=$334, median$=$321), \textit{words only} (mean$=$347, median$=$369), \textit{prosody only} (mean$=$365, median$=$369), and \textit{no words} (mean$=$319, median$=$329). The longer gaps for question transitions could give them an advantage because our anticipatory measure includes shifts initiated during the gap between turns (Figure \ref{fig:criterion}).

@@

% PART UNFINISHED: APPENDIX

>>>>

\usepackage{graphicx}
\usepackage{lineno}
\usepackage{amssymb}
\usepackage{wrapfig}
\usepackage{subfig}
\usepackage{tipa}
\usepackage{float}
\usepackage{longtable}

@@

\begin{abstract}
Children begin developing turn-taking skills in infancy but take several years to assimilate their growing knowledge of language into their turn-taking behavior. In two eye-tracking experiments, we measured children's anticipatory gaze to upcoming responders while controlling linguistic cues to upcoming turn structure. In Experiment 1, we showed English and non-English conversations to English-speaking participants, finding minimal differences between the predictive looking behavior of preschoolers and adults. In Experiment 2, we phonetically controlled lexicosyntactic and prosodic cues in English-only speech, finding that children's predictive looking behavior improved from ages one to six, but that even one-year-olds made more anticipatory looks than would be expected by chance. In both experiments, children and adults anticipated more often after hearing questions. Like adults, prosody alone did not improve childrenÕs predictive gaze shifts. But, unlike adults, lexical cues alone were also not sufficient to improve prediction---children's performance was best overall with access to lexicosyntax and prosody together. Our findings support an account in which turn prediction emerges in infancy, but takes several years before becoming fully integrated with linguistic processing. 
\end{abstract}

@@

\linenumbers

@@

Children participate in sequential coordination with their caregivers starting at three months of age---before they can rely on any linguistic cues in taking turns \citep[see, among others, ][]{bateson1975, hilbrink2015, jaffe2001, snow1977}. Of course, infant turn taking is different from adult turn taking in several ways. Infant turn taking is heavily scaffolded by caregivers, has distinct timing in comparison to adult turn taking, and lacks semantic content \citep{hilbrink2015, jaffe2001}. However, children's early, turn-structured social interactions are presumably a critical precursor to their conversational turn taking. Along these lines, early non-verbal interactions establish the protocol by which children come to use language with others. And as they acquire language, they also come to integrate it into their preverbal turn-taking systems.

@@

Conversation distinguishes itself from non-conversational turn-taking behaviors by the complexity of the turn sequencing involved. In the examples above (traffic, games, and monkeys) the set of sequence and action types is far more limited and predictable than what we find in everyday talk. For example, conversational turns come grouped into semantically-contingent sequences of action. The groups can span turn-by-turn exchanges (e.g., simple question--response, ``How are you?''--``Fine.'') or sequence-by-sequence exchanges (e.g., reciprocals, ``How are you?''--``Fine, and you?''--``Great!''). Sequences of action drive the conversation forward into the next, relevant sequences of talk (e.g., ''And you?''--``Great!''--``Why's that?''; \citealp{schegloff2007}). To take a turn, participants need to make predictions about what conversational content will be relevant next. In some cases, relevant next turns are somewhat obvious (e.g., question--response) while, in other cases, there are multiple relevant next actions to choose from or no obvious next action at all (e.g., after a closing).

@@

Adults have a lot of information at their disposal to help make accurate predictions about upcoming turn content. Lexical, syntactic, and prosodic information (e.g., \textit{wh}- words, subject-auxiliary inversion, and list intonation) can all inform addressees about upcoming linguistic structure \citep{de-ruiter2006, duncan1972, ford1996, torreira2015}. Non-verbal cues (e.g., gaze, posture, and pointing) often appear at turn-boundaries and can sometimes act as late indicators of an upcoming speaker switch \citep{rossano2009, stivers2010}. Additionally, the sequential context of a turn can make it clear what will come next: Answers after questions, thanks or denial after compliments, et cetera \citep{schegloff2007}.

Prior work suggests that adult listeners primarily use lexicosyntactic information to accurately predict upcoming turn structure \citep{de-ruiter2006}. De Ruiter and colleagues \citeyearpar{de-ruiter2006} asked participants to listen to snippets of spontaneous conversation and to press a button whenever they anticipated that the current speaker was about to finish his or her turn. The speech snippets were controlled for the amount of linguistic information present; some were normal, but others had flattened pitch, low-pass filtered speech, or further manipulations. De Ruiter and colleagues found that, with pitch-flattened speech, the timing of participants' button responses was comparable to their timing with the full linguistic signal. But when no lexical information was available, participants' responses were significantly earlier. The authors concluded that lexicosyntactic information\footnote{The ``lexicosyntactic'' condition only included flattened pitch and so was not exclusively lexicosyntactic---the speech would still have residual prosodic structure, including syllable duration and intensity.} was necessary and possibly sufficient for turn-end projection, while intonation was neither necessary nor sufficient. Congruent evidence comes from studies varying the predictability of lexicosyntactic and pragmatic content: Adults anticipate turn ends better when they can more accurately predict the exact words that will come next (\citealp{magyari2012}; see also \citealp{magyari2014}). They can also identify speech acts within the first word of an utterance \citep{gisladottir2015}, allowing them to start planning their response at the first moment possible \citep{bogels2015}.

The role of prosody for adult turn prediction is still a matter of debate. De Ruiter and colleagues' (2006) experiment focused on the role of intonation, which is only a partial index of prosody. Prosodic structure is also tied closely to the syntax of an utterance, and so the two linguistic signals are difficult to control independently \citep{ford1996}. \citet*{torreira2015} used a combination of button-press and verbal responses to investigate the relationship between lexicosyntactic and prosodic cues in turn-end prediction. Critically, their stimuli were cross-spliced so that each item had full prosodic cues to accompany the lexicosyntax. Because of the splicing, they were able to create items that had syntactically-complete units with no intonational phrase boundary at the end. Participants never verbally responded or pressed the ``turn-end'' button when hearing a syntactically-complete phrase without an intonational phrase boundary. And when intonational phrase boundaries were embedded in multi-utterance turns, participants were tricked into pressing the ``turn-end'' button 29\% of the time. Their results suggest that listeners actually do rely on prosodic cues to execute a response (see also de \citet{de-ruiter2006}:525). These experimental findings corroborate other corpus and experimental work promoting a combination of cues (lexicosyntactic, prosodic, and pragmatic) as key for accurate turn-end prediction \citep{duncan1972, ford1996, hirvenkari2013}.

@@

The timing of children's responses to their caregivers' speech shows a non-linear pattern of fall-rise-fall from early infancy to middle childhood. A recent study by Hilbrink et al. (2015) finds that infants' turn timing at three months is often too early or too late: They start vocalizing in overlap on $~$40\% of their caregivers' turns, and their non-overlapped vocalizations come after an average inter-turn silent gap of 350--900 msec (adult average: 200 msec). Between four and nine months, children begin to reduce the number of turns happening in overlap while also improving on their average response latency. But then, later on, children's response latencies slow down again, peaking at average gaps of more than 1000 msec at nine months, with only very gradual improvement after that \citep{hilbrink2015}. While children's avoidance of overlap is nearly adult-like by nine months, the timing of their non-overlapped responses stays much longer than the 200 msec standard for the next few years \citep{casillas2016, garvey1984, ervin-tripp1979}.

@@

Keitel and colleagues \citeyearpar{keitel2013} addressed the random baseline issue in their study of 6-, 12-, 24-, and 36-month-olds. They asked participants to watch short videos of conversation and tracked their eye movements at points of speaker change. They found that children's anticipatory gaze frequency was only greater than chance for 36-month-olds and adults. Their study was the first to focus on the role of linguistic processing in children's turn predictions. They showed their participants two types of conversation videos: One normal and one with flattened pitch (i.e., with flattened intonation contours), finding that only 36-month-olds were affected by a lack of intonation contours. The adult control group made equal numbers of anticipatory looks in the videos, with and without intonation contours, consistent with prior adult findings \citep{de-ruiter2006}. Keitel and colleagues concluded that children's ability to predict upcoming turn structure relies on their ability to comprehend the stimuli (emerging around 36 months), especially with respect to semantic access. They also suggest that intonation takes a secondary role in turn prediction, but only \textit{after} children acquire more sophisticated, adult-like language comprehension systems (sometime after 36 months).

Although the Keitel et al. \citeyearpar{keitel2013} study constitutes a substantial advance over previous work, it has its own limitations. Because these limitations directly inform our own study design, we review them in some detail. First, their estimates of baseline gaze frequency (``random'' in their terminology) were not random. Instead, they used gaze switches during ongoing speech as a baseline, during which switching is least likely to occur \citep{hirvenkari2013} and thereby maximizing their chances of finding a difference between gaze frequency at turn transitions and their baseline rate. A more conservative baseline would be to compare participants' looking behavior at turn transitions to their looking behavior during randomly selected windows of time throughout the stimulus, including turn transitions. We follow this conservative approach in our work.

@@

Despite these minor methodological issues, the Keitel et al. \citeyearpar{keitel2013} study still demonstrates intriguing age-based differences in children's ability to predict upcoming turn structure, and the results suggest that both semantic and intonational development \textit{do} play a role in children's looking patterns. Our current work thus takes this paradigm as a starting point.\footnote{See also \citet{casillas2012, casillas2013}.}

@@

We report here on the role of linguistic processing in children's predictions about upcoming turn structure, in particular on how children use prosodic and lexicosyntactic information to make their predictions. Prior work has focused mainly on lexicosyntax and intonation, and not on prosody proper (\citealp{de-ruiter2006}; \citealp{keitel2013}, but see \citealp{torreira2015}), even though infants seem to acquire the basic rhythmic properties of the prosodic signal first \citep{mehler1988, moon1993, nazzi2003}.

@@

Each recording session began with a 20-minute warm-up period of spontaneous conversation during which the pair talked for five minutes on four topics (favorite foods, entertainment, hometown layout, and pets). Then we asked talkers to choose a new topic---one relevant to young children (e.g., riding a bike, eating breakfast)---and to improvise a dialogue on that topic. We asked them to speak as if they were on a children's television show in order to elicit child-directed speech toward each other. We recorded until the talkers achieved at least 30 seconds of uninterrupted discourse with enthusiastic, child-directed speech. Most talker pairs took less than five minutes to complete the task, usually by agreeing on a rough script at the start. We encouraged talkers to ask at least a few questions to each other during the improvisation. The resulting conversations were therefore not entirely spontaneous, but were as close as possible while still remaining child-oriented in topic, prosodic pattern, and lexicosyntactic construction.\footnote{All of the non-English talkers were fluent in English as a second language, and some fluently spoke three or more languages. We chose male-female pairs as a natural way of creating contrast between the two talker voices.}

@@

Because the conversational stimuli were recorded semi-spontaneously, the duration of turn transitions and the number of speaker transitions in each video was variable. We measured the duration of each turn transition from the audio recording associated with each video. We excluded turn transitions longer than 550 msec and shorter than 90 msec, including overlapped transitions, from analysis.\footnote{Overlap occurs when a responder begins a new turn before the current turn is finished. When overlap occurs, observers cannot switch their gaze in anticipation of the response because the response began earlier than expected; participants expect conversations to proceed with ``one speaker at a time'' \citep{sacks1974}. As such, they would still be fixated on the prior speaker when the overlap started, and then would have to switch their gaze \textit{reactively} to the responder.} This left approximately equal numbers of turn transitions available for analysis in the English (N$=$20) and non-English (N$=$16) videos. On average, the inter-turn gaps for English videos (mean$=$318, median$=$302, stdev$=$112 msec) were slightly longer than for non-English videos (mean$=$286, median$=$251, stdev$=$122 msec). The longer gaps in the English videos could give them a slight advantage: Our definition of an ``anticipatory gaze shift'' includes shifts that are initiated during the gap between turns (Figure \ref{fig:criterion}), so participants had slightly more time to make anticipatory shifts in the English videos.

Questions made up exactly half of the turn transitions in the English (N$=$10) and non-English (N$=$8) videos. In the English videos, inter-turn gaps were slightly shorter for questions (mean$=$310, median$=$293, stdev$=$112 msec) than non-questions (mean$=$325, median$=$315, stdev$=$118 msec). Non-English videos did not show a large difference in transition time for questions (mean$=$270, median$=$257, stdev$=$116 msec) and non-questions (mean$=$302, median$=$252, stdev$=$134 msec).

@@

\begin{table}[h!]
  \begin{scriptsize}
\begin{center}
  \begin{tabular}{lcccl}
  \textbf{\textit{Children}} &&&& \\
    \hline
                   & Estimate & Std. Error & z value & Pr($>$$|$z$|$)    \\
    \hline
    (Intercept)   &      -0.96146 &   0.84901 & -1.132 & 0.257446 \\    
    Age             &    -0.18268 &   0.17507 & -1.043 & 0.296725   \\  
    LgCond$=$\textit{non-English} &          -3.29347 &   0.96045 & -3.429 & 0.000606 *** \\
    Type$=$\textit{non-Question}          &     -1.10129 &   0.86494 & -1.273 & 0.202925     \\
    Duration      &       3.40169 &   1.22826 &  2.770 & 0.005614  **\\
    Age*LgCond$=$\textit{non-English} &       0.52065 &   0.21190 &  2.457 & 0.014008 ** \\  
    Age*TypeS$=$\textit{non-Question}          & -0.01628 &   0.19437 & -0.084 & 0.933232     \\
    LgCond$=$\textit{non-English}* &     2.68166 &   1.35016 &  1.986 & 0.047013 * \\
    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
    Age*LgCond$=$\textit{non-English}* & -0.45632 &   0.30163 & -1.513 & 0.130315  \\
    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
    \hline
  &&&& \\
  \textbf{\textit{Adults}} &&&& \\
    \hline
                       &  Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\    
    \hline
    (Intercept)       &        -0.1966 &    0.6942 & -0.283 & 0.776988     \\
    LgCond$=$\textit{non-English}    &             -0.8812 &    0.9602 & -0.918 & 0.358754   \\  
    Type$=$\textit{non-Question}             &        -4.4953 &    1.3139 & -3.421 & 0.000623 *** \\
    Duration          &        -1.1227 &    1.9880 & -0.565 & 0.572238     \\
    LgCond$=$\textit{non-English}* &           3.2972 &    1.6101 &  2.048 & 0.040581 * \\ 
    Type$=$\textit{non-Question} &&&& \\
    LgCond$=$\textit{non-English}* &        1.3626 &    3.0077 &  0.453 & 0.650527     \\
    \hspace*{5mm} Duration &&&& \\
    Type$=$\textit{non-Question}*          &  10.5107 &    3.3459 &  3.141 & 0.001682  ** \\
    \hspace*{5mm} Duration &&&& \\
    LgCond$=$\textit{non-English}* & -6.3156 &    4.4926 & -1.406 & 0.159790 \\    
    Type$=$\textit{non-Question}* &&&& \\
    \hspace*{5mm} Duration &&&& \\
    \hline
  \end{tabular}
\end{center}
  \end{scriptsize}
  \caption{Model output for children and adults' anticipatory gaze switches.}
\label{tab:E1-models}
\end{table}

@@

Our baseline analyses revealed that none of the significant predictors from models of the original, turn-related data can be explained by random looking. The children's data showed strong evidence of differentiation from the randomly permuted data for all four significant effects in the original model (Table \ref{tab:E1-models}: Children):  the original \textit{t}-values for language condition, gap duration, the age-language condition interaction, and the language condition-transition type interaction were all greater than 95\% of \textit{t}-values for the randomly permuted data (99.9\%, 95.5\%, 99.4\%, and 96\%, respectively; Figure \ref{fig:E1-ChiTs}). Similarly, the adults' data showed significant differentiation from the randomly permuted data for two of the three originally significant predictors---transition type and the transition type-gap duration interaction (greater than 99.9\% and 99.7\% of random \textit{t}-values, respectively)---with marginal differentiation for the interaction of language condition and transition type (greater than 94.7\% of random \textit{t}-values; Figure \ref{fig:E1-AduTs}). The effects of language condition and transition type for the real and randomly permuted data can also be observed in Figure \ref{fig:E1-randvsreal}. We excluded the output of random-permutation models that did not ultimately converge to remove unreliable model results from our percentile calculations below (78\% and 76\% of models for children and adults, respectively).

@@

@@

There were 85 turn transitions for analysis after excluding transitions longer than 550 msec and shorter than 90 msec. The remaining turn transitions had slightly more questions than non-question (N$=$50 and N$=$35, respectively), with transitions distributed somewhat evenly across conditions (keeping in mind that there were three \textit{normal} videos and only one limited cue video for each experiment version): \textit{Normal} (N$=$36), \textit{words only} (N$=$13), \textit{prosody only} (N$=$17), and {no speech} (N$=$19). Inter-turn gaps for questions (mean$=$365, median$=$427) were longer than those for non-questions (mean$=$302, median$=$323) on average, but gap duration was overall comparable across conditions: \textit{Normal} (mean$=$334, median$=$321), \textit{words only} (mean$=$347, median$=$369), \textit{prosody only} (mean$=$365, median$=$369), and \textit{no words} (mean$=$319, median$=$329). The longer gaps for question transitions could give them an advantage because our anticipatory measure includes shifts initiated during the gap between turns (Figure \ref{fig:criterion}).

@@

\appendix
\setcounter{figure}{0}  
In all of the following plots, the gray dots represent the randomly permuted data's model estimates for the value listed (beta or standard error), the white dots represent the model estimates from the original data, and the triangles represent the 95th percentile for each distribution being shown.

\begin{figure}
  \centering
  \subfloat[Children]{\includegraphics[width=0.49\textwidth]{figures/E1-chi-randrun-betas-absolute.png}\label{fig:E1-ChiBs}}
  \hfill
  \subfloat[Adults]{\includegraphics[width=0.49\textwidth]{figures/E1-adu-randrun-betas-absolute.png}\label{fig:E1-AduBs}}
  \caption{Random-permutation and original $|$\textit{$\beta$}-values$|$ for predictors of children and adults' anticipatory gaze rates in Experiment 1.}
\end{figure}

\begin{figure}
  \centering
  \subfloat[Children]{\includegraphics[width=0.49\textwidth]{figures/E1-chi-randrun-SEs-absolute.png}\label{fig:E1-ChiSEs}}
  \hfill
  \subfloat[Adults]{\includegraphics[width=0.49\textwidth]{figures/E1-adu-randrun-SEs-absolute.png}\label{fig:E1-AduSEs}}
  \caption{Random-permutation and original $|$\textit{SE}-values$|$ for predictors of children and adults' anticipatory gaze rates in Experiment 1.}
\end{figure}
